---
title: "Homework 2 - Divij Sanjanwala"
output: html_document
always_allow_html: yes
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, message = FALSE, warning = FALSE)
```

```{r}
## R packages 
setwd(getwd())
library(SpatialEpi)
library(tidyverse)
library(sf)
library(spdep)
library(rmarkdown)
library(INLA)
library(tmap) ## Very commonly used -- to make static and interactive plots
library(mapview)
library(cowplot)
library(reshape2)
library(leaflet)
library(raster)
library(shiny)
library(tidyverse)
library(spDataLarge) ## Data sets -- use commands in notes to download
library(spData) ## Data sets
library(sf) ## simple features 
```
# Question 1.1:

## For the three following sets of prior distributions, generate 20 data sets from the respective prior predictive distribution.
```{r}
#-----------------------------------------------
# Setting the values of the parameters
#-----------------------------------------------
beta0 <- 1
beta1 <- 0.5
sigma <- 1
#-----------------------------------------------
# Simulating covariate values + data
#-----------------------------------------------
set.seed(17)
x <- runif(n = 100, min = 1, max=5)
y.mean <- beta0 + beta1*x
y <- rnorm(n = 100, 
           mean = y.mean, 
           sd = sigma)
sim.data <- tibble(x,y, y.mean)
#-----------------------------------------------
```

- Simulation for prior distribution candidate:
$$\begin{equation}
\beta_0 \sim N\left(0, \sigma_{\beta_0}=1\right), \beta_1 \sim N\left(0, \sigma_{\beta_1}=1\right), \sigma \sim \text { Gamma }(\text { shape }=1, \text { scale }=1)
\end{equation}$$

```{r}
set.seed(17)
n <- 100

sigma <- rgamma(1, shape = 1, scale = 1)
beta0 <- rnorm(1, mean = 0, sd = 1)
beta1 <- rnorm(1, mean = 0, sd = 1)

x <- runif(n = 100, min = 1, max=5)
y.mean <- beta0 + beta1*x

sim.data20_1 <- tibble(x = rep(x, 20), 
              y.mean = rep(y.mean, 20), 
              y.sim = rnorm(n = 20*100, 
                            mean = y.mean, 
                            sd = sigma), 
              group = rep(1:20, each = 100))

ggplot(data = sim.data20_1, aes(x,y.sim)) + 
  geom_point() + 
  geom_line(aes(x,y.mean), col="blue") + 
  ylab("Simulated Data") + 
  facet_wrap(~group) +
  theme_bw()
```

```{r}
beta0_1 = mean(dnorm(0, 1))
beta1_1 = mean(dnorm(0, 1))
sigma_1 = mean(dgamma(1, 1))

ggplot(data.frame(bx = seq(-12, 12, length.out = 100)), aes(bx)) +
  stat_function(fun = dnorm, args = list(mean = 0, sd = 1), aes(color = "Beta_0")) + 
  stat_function(fun = dnorm, args = list(mean = 0, sd = 1), aes(color = "Beta_1")) + 
  stat_function(fun = dgamma, args=list(shape = 1, scale =1), aes(color = "Sigma")) +
  geom_vline(xintercept = beta0_1, linetype = "dashed", colour = "red") +
  geom_vline(xintercept = beta1_1, linetype = "dashed", colour = "blue") +
  geom_vline(xintercept = sigma_1, linetype = "dashed", colour = "green") +
  ylab("density") + 
  xlab("x") + 
  scale_color_manual(values = c("red", "blue", "green"), 
                     labels = c("Beta_0", "Beta_1", "Sigma")) +
  labs(title = "Prior distributions for Beta_0, Beta_1, and Sigma",
       x = "Prior Value",
       y = "Density",
       color = "Priors")
```

- Simulation for prior distribution candidate:
$$\begin{equation}
\beta_0 \sim N\left(0, \sigma_{\beta_0}=1000\right), \beta_1 \sim N\left(0, \sigma_{\beta_1}=1000\right), \sigma \sim \text { Gamma }(\text {shape}=1000, \text {scale}=1000)
\end{equation}$$

```{r}
set.seed(17)
n <- 100

sigma <- rgamma(1, shape = 1000, scale = 1000)
beta0 <- rnorm(1, mean = 0, sd = 1000)
beta1 <- rnorm(1, mean = 0, sd = 1000)

x <- runif(n = 100, min = 1, max=5)
y.mean <- beta0 + beta1*x

sim.data20_2 <- tibble(x = rep(x, 20), 
                       y.mean = rep(y.mean, 20), 
                       y.sim = rnorm(n = 20*100, mean = y.mean, sd = sigma), 
                       group = rep(1:20, each = 100))

ggplot(data = sim.data20_2, aes(x,y.sim)) + 
  geom_point() + 
  geom_line(aes(x,y.mean), col="blue") + 
  ylab("Simulated Data") + 
  facet_wrap(~group) +
  theme_bw()
```

```{r}
beta0_2 = mean(dnorm(0, 1000))
beta1_2 = mean(dnorm(0, 1000))
sigma_2 = mean(dgamma(1000, 1000))

ggplot(data.frame(bx = seq(-100000, 100000, length.out = 100)), aes(bx)) +
  stat_function(fun = dnorm, args = list(mean = 0, sd = 1000), aes(color = "Beta_0")) + 
  stat_function(fun = dnorm, args = list(mean = 0, sd = 1000), aes(color = "Beta_1")) + 
  stat_function(fun = dgamma, args=list(shape = 1000, scale =1000), aes(color = "Sigma")) +
  geom_vline(xintercept = beta0_2, linetype = "dashed", colour = "red") +
  geom_vline(xintercept = beta1_2, linetype = "dashed", colour = "blue") +
  geom_vline(xintercept = sigma_2, linetype = "dashed", colour = "green") +
  ylab("density") + 
  xlab("x") + 
  scale_color_manual(values = c("red", "blue", "green"), 
                     labels = c("Beta_0", "Beta_1", "Sigma")) +
  labs(title = "Prior distributions for Beta_0, Beta_1, and Sigma",
       x = "Prior Value",
       y = "Density",
       color = "Priors")
```

- Simulation for prior distribution candidate:
$$\begin{equation}
\beta_0 \sim Unif\left(0, 1\right), \beta_1 \sim Unif\left(-1, 0\right), \sigma \sim \text { Exp }(1)
\end{equation}$$

```{r}
set.seed(17)
n <- 100
data_sets <- list()

sigma <- rexp(1, rate = 1)
beta0 <- runif(1, min = 0, max = 1)
beta1 <- runif(1, min = -1, max = 0)
  
x <- runif(n = 100, min = 1, max=5)
y.mean <- beta0 + beta1*x

sim.data20_3 <- tibble(x = rep(x, 20), 
              y.mean = rep(y.mean, 20), 
              y.sim = rnorm(n = 20*100, 
                            mean = y.mean, 
                            sd = sigma), 
              group = rep(1:20, each = 100))


ggplot(data = sim.data20_3, aes(x,y.sim)) + 
  geom_point() + 
  geom_line(aes(x,y.mean), col="blue") + 
  ylab("Simulated Data") + 
  facet_wrap(~group) +
  theme_bw()
```


```{r}
beta0_3 = mean(dunif(0, 1))
beta1_3 = mean(dunif(-1, 0))
sigma_3 = mean(dexp(1))

ggplot(data.frame(bx = seq(-12, 12, length.out = 100)), aes(bx)) +
  stat_function(fun = dunif, args = list(min = 0, max = 1), aes(color = "Beta_0")) + 
  stat_function(fun = dunif, args = list(min = -1, max = 0), aes(color = "Beta_1")) + 
  stat_function(fun = dexp, args=list(rate = 1), aes(color = "Sigma")) +
  geom_vline(xintercept = beta0_3, linetype = "dashed", colour = "red") +
  geom_vline(xintercept = beta1_3, linetype = "dashed", colour = "blue") +
  geom_vline(xintercept = sigma_3, linetype = "dashed", colour = "green") +
  ylab("density") + 
  xlab("x") + 
  scale_color_manual(values = c("red", "blue", "green"), 
                     labels = c("Beta_0", "Beta_1", "Sigma")) +
  labs(title = "Prior distributions for Beta_0, Beta_1, and Sigma",
       x = "Prior Value",
       y = "Density",
       color = "Priors")
  
```
The best choice of the betas and sigma combinations would be to consider:

$$\begin{equation}
\beta_0 \sim N\left(0, \sigma_{\beta_0}=1\right), \beta_1 \sim N\left(0, \sigma_{\beta_1}=1\right), \sigma \sim \text { Gamma }(\text { shape }=1, \text { scale }=1)
\end{equation}$$

This is because, simulated plots seem to uniformly fit the mean line and have a clear pattern across the datasets. Considering the density plots as well, they appear to be close to the true value of the mean. This translates into effective simulations.
# Question 1.2:

```{r}
#-----------------------------------------------
# Setting the values of the parameters
#-----------------------------------------------
set.seed(17)
nu.mu <- 2
tau.mu <- 0.5
nu.beta <- -1
tau.beta <- 0.5
mu.hm <- rnorm(n=20, mean = nu.mu, sd=tau.mu)
beta.hm <- rnorm(n=20, mean = nu.beta, sd= tau.beta)
sigma <- 1
#-----------------------------------------------
# Simulating covariate values + data
#-----------------------------------------------
x.hm <- runif(n = 100, min = 1, max=5)
y.mean.hier <- c(rep(mu.hm, each = 100) + 
                   rep(beta.hm, each = 100)*rep(x.hm, 20))

y.hier <- rnorm(n = 20*100, mean = y.mean.hier, sigma)
sim.data.hier <- tibble(x = rep(x.hm, 20), 
                        y.hier, y.mean.hier, 
                        group = paste("Group", rep(1:20, each = 100)))
```

- Simulation for prior distribution candidate:

$$
\begin{aligned}
& \nu_\mu \sim N(0,1), \nu_\beta \sim N(0,1), \tau_\mu \sim \text { Gamma }(\text { shape }=1, \text { scale }=1), \tau_\beta \sim \text { Gamma }(\text { shape }=1, \text { scale }= \\
& 1), \sigma \sim \text { Gamma }(\text { shape }=1, \text { scale }=1)
\end{aligned}
$$

```{r}
#-----------------------------------------------
# Setting the values of the parameters
#-----------------------------------------------
set.seed(17)
nu.mu <- rnorm(1, mean=0, sd=1)
tau.mu <- rgamma(1, shape=1, scale=1)
nu.beta <- rnorm(1, mean=0, sd=1)
tau.beta <- rgamma(1, shape=1, scale=1)
mu.hm <- rnorm(n=20, mean = nu.mu, sd=tau.mu)
beta.hm <- rnorm(n=20, mean = nu.beta, sd= tau.beta)
sigma <- rgamma(1, shape=1, scale=1)
#-----------------------------------------------
# Simulating covariate values + data
#-----------------------------------------------
x.hm <- runif(n = 100, min = 1, max=5)
y.mean.hier <- c(rep(mu.hm, each = 100) + 
                   rep(beta.hm, each = 100)*rep(x.hm, 20))

y.hier <- rnorm(n = 20*100, mean = y.mean.hier, sigma)
sim.data.hier <- tibble(x = rep(x.hm, 20), 
                        y.hier, y.mean.hier, 
                        group = paste("Group", rep(1:20, each = 100)))

ggplot(sim.data.hier, aes(x, y.hier)) + 
  geom_point() + geom_line(aes(x, y.mean.hier), col="blue") + 
  facet_wrap(~group)
```
```{r}

nu.mu <- dnorm(1, mean=0, sd=1)
tau.mu <- dgamma(1, shape=1, scale=1)
nu.beta <- dnorm(1, mean=0, sd=1)
tau.beta <- dgamma(1, shape=1, scale=1)
sigma <- dgamma(1, shape=1, scale=1)

ggplot(data.frame(bx = seq(-12, 12, length.out = 100)), aes(bx)) +
  stat_function(fun = dnorm, args = list(mean = 0, sd = 1), aes(color = "nu.mu")) + 
  stat_function(fun = dgamma, args = list(shape = 1, scale = 1), aes(color = "tau.mu")) + 
  stat_function(fun = dgamma, args = list(shape = 1, scale = 1), aes(color = "tau.beta")) + 
  stat_function(fun = dnorm, args = list(mean=0, sd=1), aes(color = "nu.beta")) + 
  stat_function(fun = dgamma, args=list(shape = 1, scale = 1), aes(color = "Sigma")) +
  geom_vline(xintercept = nu.mu, linetype = "dashed", colour = "red") +
  geom_vline(xintercept = tau.mu, linetype = "dashed", colour = "blue") +
  geom_vline(xintercept = nu.beta, linetype = "dashed", colour = "green") +
  geom_vline(xintercept = tau.beta, linetype = "dashed", colour = "pink") +
  geom_vline(xintercept = sigma, linetype = "dashed", colour = "purple") +
  ylab("density") + 
  xlab("x") + 
  scale_color_manual(values = c("red", "blue", "green", "pink", "black", "purple"), 
                     labels = c("nu.mu", "tau.mu", "tau.beta", "nu.beta", "tau.mu")) +
  labs(title = "Prior distributions for nu.mu, tau.mu, tau.beta, nu.beta, tau.mu",
       x = "Prior Value",
       y = "Density",
       color = "Priors")

```
- Simulation for prior distribution candidate:

$$
\begin{aligned}
& \nu_\mu \sim N(0,1000), \nu_\beta \sim N(0,1000), \tau_\mu \sim \text { Gamma }(\text { shape }=1000, \text { scale }=1000), \tau_\beta \sim \text { Gamma }(\text { shape }=1, \text { scale }= \\
& 1000), \sigma \sim \text { Gamma }(\text { shape }=1000, \text { scale }=1000)
\end{aligned}
$$

```{r}
#-----------------------------------------------
# Setting the values of the parameters
#-----------------------------------------------
set.seed(17)
nu.mu <- rnorm(1, mean=0, sd=1000)
tau.mu <- rgamma(1, shape=1000, scale=1000)
nu.beta <- rnorm(1, mean=0, sd=1000)
tau.beta <- rgamma(1, shape=1, scale=1000)
mu.hm <- rnorm(n=20, mean = nu.mu, sd=tau.mu)
beta.hm <- rnorm(n=20, mean = nu.beta, sd= tau.beta)
sigma <- rgamma(1, shape=1000, scale=1000)
#-----------------------------------------------
# Simulating covariate values + data
#-----------------------------------------------
x.hm <- runif(n = 100, min = 1, max=5)
y.mean.hier <- c(rep(mu.hm, each = 100) + 
                   rep(beta.hm, each = 100)*rep(x.hm, 20))

y.hier <- rnorm(n = 20*100, mean = y.mean.hier, sigma)
sim.data.hier <- tibble(x = rep(x.hm, 20), 
                        y.hier, y.mean.hier, 
                        group = paste("Group", rep(1:20, each = 100)))

ggplot(sim.data.hier, aes(x, y.hier)) + 
  geom_point() + geom_line(aes(x, y.mean.hier), col="blue") + 
  facet_wrap(~group)
```

```{r}
nu.mu <- rnorm(1, mean=0, sd=1000)
tau.mu <- rgamma(1, shape=1000, scale=1000)
nu.beta <- rnorm(1, mean=0, sd=1000)
tau.beta <- rgamma(1, shape=1, scale=1000)
sigma <- rgamma(1, shape=1000, scale=1000)

ggplot(data.frame(bx = seq(-12, 12, length.out = 100)), aes(bx)) +
  stat_function(fun = dnorm, args = list(mean = 0, sd = 1000), aes(color = "nu.mu")) + 
  stat_function(fun = dgamma, args = list(shape = 1000, scale = 1000), aes(color = "tau.mu")) + 
  stat_function(fun = dgamma, args = list(shape = 1000, scale = 1000), aes(color = "tau.beta")) + 
  stat_function(fun = dnorm, args = list(mean=0, sd=1000), aes(color = "nu.beta")) + 
  stat_function(fun = dgamma, args=list(shape = 1000, scale = 1000), aes(color = "Sigma")) +
  geom_vline(xintercept = nu.mu, linetype = "dashed", colour = "red") +
  geom_vline(xintercept = tau.mu, linetype = "dashed", colour = "blue") +
  geom_vline(xintercept = nu.beta, linetype = "dashed", colour = "green") +
  geom_vline(xintercept = tau.beta, linetype = "dashed", colour = "pink") +
  geom_vline(xintercept = sigma, linetype = "dashed", colour = "purple") +
  ylab("density") + 
  xlab("x") + 
  scale_color_manual(values = c("red", "blue", "green", "pink", "black", "purple"), 
                     labels = c("nu.mu", "tau.mu", "tau.beta", "nu.beta", "tau.mu")) +
  labs(title = "Prior distributions for nu.mu, tau.mu, tau.beta, nu.beta, tau.mu",
       x = "Prior Value",
       y = "Density",
       color = "Priors")
```

- Simulation for prior distribution candidate:

$$
\begin{aligned}
& \nu_\mu \sim Unif(0,1), \nu_\beta \sim Unif(0,1), \tau_\mu \sim \text { Exp }(1), \tau_\beta \sim \text { Unif }(1), \sigma \sim \text { Unif }(1)
\end{aligned}
$$

```{r}
#-----------------------------------------------
# Setting the values of the parameters
#-----------------------------------------------
set.seed(17)
nu.mu <- runif(1, min=0, max=1)
tau.mu <- rexp(1, rate=1)
nu.beta <- runif(1, min=0, max=1)
tau.beta <- rexp(1, rate=1)
mu.hm <- rnorm(n=20, mean = nu.mu, sd=tau.mu)
beta.hm <- rnorm(n=20, mean = nu.beta, sd= tau.beta)
sigma <- rexp(1, rate=1)
#-----------------------------------------------
# Simulating covariate values + data
#-----------------------------------------------
x.hm <- runif(n = 100, min = 1, max=5)
y.mean.hier <- c(rep(mu.hm, each = 100) + 
                   rep(beta.hm, each = 100)*rep(x.hm, 20))

y.hier <- rnorm(n = 20*100, mean = y.mean.hier, sigma)
sim.data.hier <- tibble(x = rep(x.hm, 20), 
                        y.hier, y.mean.hier, 
                        group = paste("Group", rep(1:20, each = 100)))

ggplot(sim.data.hier, aes(x, y.hier)) + 
  geom_point() + geom_line(aes(x, y.mean.hier), col="blue") + 
  facet_wrap(~group)
```

```{r}

nu.mu <- runif(1, min=0, max=1)
tau.mu <- rexp(1, rate=1)
nu.beta <- runif(1, min=0, max=1)
tau.beta <- rexp(1, rate=1)
sigma <- rexp(1, rate=1)

ggplot(data.frame(bx = seq(-12, 12, length.out = 100)), aes(bx)) +
  stat_function(fun = dunif, args = list(min = 0, max = 1), aes(color = "nu.mu")) + 
  stat_function(fun = dexp, args = list(1), aes(color = "tau.mu")) + 
  stat_function(fun = dunif, args = list(min = 0, max = 1), aes(color = "nu.beta")) + 
  stat_function(fun = dexp, args = list(1), aes(color = "tau.beta")) + 
  stat_function(fun = dexp, args=list(1), aes(color = "Sigma")) +
  geom_vline(xintercept = nu.mu, linetype = "dashed", colour = "red") +
  geom_vline(xintercept = tau.mu, linetype = "dashed", colour = "blue") +
  geom_vline(xintercept = nu.beta, linetype = "dashed", colour = "green") +
  geom_vline(xintercept = tau.beta, linetype = "dashed", colour = "pink") +
  geom_vline(xintercept = sigma, linetype = "dashed", colour = "purple") +
  ylab("density") + 
  xlab("x") + 
  scale_color_manual(values = c("red", "blue", "green", "pink", "black", "purple"), 
                     labels = c("nu.mu", "tau.mu", "tau.beta", "nu.beta", "tau.mu")) +
  labs(title = "Prior distributions for nu.mu, tau.mu, tau.beta, nu.beta, tau.mu",
       x = "Prior Value",
       y = "Density",
       color = "Priors")





```
The best choice of the betas and sigma combinations would be to consider:

$$\begin{equation}
\beta_0 \sim N\left(0, \sigma_{\beta_0}=1\right), \beta_1 \sim N\left(0, \sigma_{\beta_1}=1\right), \sigma \sim \text { Gamma }(\text { shape }=1, \text { scale }=1)
\end{equation}$$

This is because, simulated plots seem to uniformly fit the mean line and have a clear pattern across the datasets. Considering the density plots as well, they appear to be close to the true value of the mean. This translates into effective simulations.
# Question 2.1: Fitting a Linear Regression Model

```{r}
load("bayes-vis.RData")
## Adding a column with numbers
GM$super_region_name_dup <- GM$super_region_name
latcab <- st_as_sf(GM[GM$super_region == 5,])
```

## First set of priors for linear regression:

- $\beta_0 \sim \mathrm{N}\left(\mu=0, \sigma^2=1 / 0.1\right)$
- $\beta_1 \sim N\left(\mu=0, \sigma^2=1 / 0.1\right)$
- $\tau \sim \operatorname{Gamma}(0.01,0.01)$

```{r}
## Adding in prior distributions
### Default normal distributions for b0 and b1
prior.fixed1_model1 <- list(mean.intercept = 0, prec.intercept = 0.1,
                    mean = 0, prec = 0.1)
prior.prec1_model1 <- list(prec = list(prior = "loggamma", 
                               param = c(0.01, 0.01)))

completepool1_model1 <- inla(formula = pm25 ~ 1 + sat_2014, 
                     data = data.frame(latcab), 
                     control.fixed = prior.fixed1_model1, 
                     control.family = list(hyper = list(prec = prior.prec1_model1)),
                     control.compute=list(config = TRUE))

postdraws.cpool1_model1 <- inla.posterior.sample(1000, completepool1_model1)

# make a dataframe of the posterior draws from 104 to 110
posterior1_model1 <- data.frame()
for(j in 1:1000){
  posterior1_model1 <- rbind(posterior1_model1, postdraws.cpool1_model1[[j]]$latent[104:105])
  colnames(posterior1_model1) <- c("Intercept", "sat_2014")
}

# Create a grid of histograms and density plots
ggplot(melt(posterior1_model1), aes(x=value)) +
  geom_histogram(aes(y=..density..)) +
  geom_density() +
  facet_wrap(~variable, scales="free") +
  theme_bw()
```
```{r}
summary(posterior1_model1)
```
## Second set of priors for linear regression:

- $\beta_0 \sim \mathrm{N}\left(\mu=0, \sigma^2=1 / 0.01\right)$
- $\beta_1 \sim N\left(\mu=0, \sigma^2=1 / 0.01\right)$
- $\tau \sim \operatorname{Gamma}(1,1)$

```{r}
## Adding in prior distributions
### Default normal distributions for b0 and b1
prior.fixed1_model2 <- list(mean.intercept = 0, prec.intercept = 0.01,
                    mean = 0, prec = 0.01)
prior.prec1_model2 <- list(prec = list(prior = "loggamma", 
                               param = c(1, 1)))

completepool1_model2 <- inla(formula = pm25 ~ 1 + sat_2014, 
                     data = data.frame(latcab), 
                     control.fixed = prior.fixed1_model2, 
                     control.family = list(hyper = list(prec = prior.prec1_model2)),
                     control.compute=list(config = TRUE))

postdraws.cpool1_model2 <- inla.posterior.sample(1000, completepool1_model2)
# make a dataframe of the posterior draws from 104 to 110
posterior1_model2 <- data.frame()
for(j in 1:1000){
  posterior1_model2 <- rbind(posterior1_model2, postdraws.cpool1_model2[[j]]$latent[104:105])
  colnames(posterior1_model2) <- c("Intercept", "sat_2014")
}

# Create a grid of histograms and density plots
ggplot(melt(posterior1_model2), aes(x=value)) +
  geom_histogram(aes(y=..density..)) +
  geom_density() +
  facet_wrap(~variable, scales="free") +
  theme_bw()

```
```{r}
summary(posterior1_model2)
```
# Question 2.2: Fitting a Multilevel Regression Model

## First set of priors for multilevel:

- $\beta_0 \sim \mathrm{N}\left(\mu=0, \sigma^2=1 / 0.1\right)$
- $\beta_1 \sim N\left(\mu=0, \sigma^2=1 / 0.1\right)$
- $\tau \sim \operatorname{Gamma}(0.01,0.01)$
- $\tau_\mu \sim \operatorname{Gamma}(0.01,0.01)$
- $\tau_\beta \sim \operatorname{Gamma}(0.01,0.01)$

```{r}
### Default normal distributions for b0 and b1
prior.fixed2_model1 <- list(mean.intercept = 0, prec.intercept = 0.1,
                    mean = 0, prec = 0.1)
prior.prec2_model1 <- list(prec = list(prior = "loggamma", 
                               param = c(0.01, 0.01)))

partialpool2_model1 <- inla(formula = pm25 ~ 1 +    
                       f(super_region_name,  
                         model = "iid", 
                         hyper = prior.prec2_model1) + 
                       f(super_region_name_dup, sat_2014, 
                         model = "iid",
                         hyper = prior.prec2_model1), 
                     data = data.frame(GM), 
                     control.fixed = prior.fixed2_model1,
                     control.family = list(hyper = list(prec = prior.prec2_model1)), 
                    control.compute = list(config = TRUE))

postdraws.cpool2_model1 <- inla.posterior.sample(1000, partialpool2_model1)

# make a dataframe of the posterior draws from 104 to 110
posterior2_model1 <- data.frame()
for(j in 1:1000){
  posterior2_model1 <- rbind(posterior2_model1, postdraws.cpool2_model1[[j]]$latent[104:110])
  colnames(posterior2_model1) <- c("super_region_name:1", "super_region_name:2", "super_region_name:3", "super_region_name:4", "super_region_name:5", "super_region_name:6", "super_region_name:7")
}

# Create a grid of histograms and density plots
ggplot(melt(posterior2_model1), aes(x=value)) +
  geom_histogram(aes(y=..density..)) +
  geom_density() +
  facet_wrap(~variable, scales="free") +
  theme_bw()
```

```{r}
summary(posterior2_model1)
```
## Second set of priors:

- $\beta_0 \sim \mathrm{N}\left(\mu=0, \sigma^2=1 / 0.01\right)$
- $\beta_1 \sim N\left(\mu=0, \sigma^2=1 / 0.01\right)$
- $\tau \sim \operatorname{Gamma}(1,1)$
- $\tau_\mu \sim \operatorname{Gamma}(1,1)$
- $\tau_\beta \sim \operatorname{Gamma}(1,1)$

```{r}
### Default normal distributions for b0 and b1
prior.fixed2_model2 <- list(mean.intercept = 0, prec.intercept = 0.01,
                    mean = 0, prec = 0.01)
prior.prec2_model2 <- list(prec = list(prior = "loggamma", 
                               param = c(1, 1)))

partialpool2_model1 <- inla(formula = pm25 ~ 1 + 
                              f(super_region_name, 
                                model = "iid", 
                                hyper = prior.prec2_model2) + 
                              f(super_region_name_dup, sat_2014, 
                                model = "iid",
                                hyper = prior.prec2_model2), 
                            data = data.frame(latcab),
                            control.fixed = prior.fixed2_model2,
                            control.family = list(hyper = list(prec = prior.prec2_model2)),
                            control.compute=list(config = TRUE))

postdraws.cpool2_model2 <- inla.posterior.sample(1000, partialpool2_model1)

# make a dataframe of the posterior draws from 104 to 110
posterior2_model2 <- data.frame()
for(j in 1:1000){
  posterior2_model2 <- rbind(posterior2_model2, postdraws.cpool2_model2[[j]]$latent[104:110])
  colnames(posterior2_model2) <- c("super_region_name:1", "super_region_name:2", "super_region_name:3", "super_region_name:4", "super_region_name:5", "super_region_name:6", "super_region_name:7")
}

# Create a grid of histograms and density plots
ggplot(melt(posterior2_model2), aes(x=value)) +
  geom_histogram(aes(y=..density..)) +
  geom_density() +
  facet_wrap(~variable, scales="free") +
  theme_bw()
```


```{r}
summary(posterior2_model2)
```

# Question 3.1:

```{r}
postdraws.cpool1_model1 <- inla.posterior.sample(100, completepool1_model1)
postdraws.cpool2_model2 <- inla.posterior.sample(100, completepool1_model2)

samples_1 <- data.frame()
for(j in 1:100){
  samples_1 <- rbind(samples_1, postdraws.cpool1_model1[[j]]$latent[104:105])
  colnames(samples_1) <- c("Intercept", "sat_2014")
}


beta0_sample_1 <- rnorm(100, mean=0, sd = 1)
beta1_sample_1 <- rnorm(100, mean=0, sd = 1)
tau1_sample_1 <- rgamma(100, shape = 0.01, scale = 0.01)

y_1_simulation <- beta0_sample_1 + samples_1$sat_2014*beta1_sample_1 + samples_1$Intercept + tau1_sample_1
# Create a grid of histograms and density plots

samples_2 <- data.frame()
for(j in 1:100){
  samples_2 <- rbind(samples_2, postdraws.cpool1_model2[[j]]$latent[104:105])
  colnames(samples_2) <- c("Intercept", "sat_2014")
}

beta0_sample_2 <- rnorm(100, mean=0, sd = 1)
beta1_sample_2 <- rnorm(100, mean=0, sd = 1)
tau1_sample_2 <- rgamma(100, shape = 0.01, scale = 0.01)

y_2_simulation <- beta0_sample_2 + samples_2$sat_2014*beta1_sample_2 + samples_2$Intercept + tau1_sample_2

plot1 <- ggplot(data.frame(y_1_simulation), aes(x=y_1_simulation)) +
  geom_histogram(aes(y=..density..)) +
  geom_density() +
  theme_bw()

plot2 <- ggplot(data.frame(y_2_simulation), aes(x=y_2_simulation)) +
  geom_histogram(aes(y=..density..)) +
  geom_density() +
  theme_bw()


plot_grid(plot1, plot2, ncol = 2)
```
The simulated data appears to to near about the same as the actual data shown in the Q2.2 histograms. Once we multiply the predictors with the hyperparameter estimates, we roughly get a similar histogram for both the models.

# Question 3.2:

```{r}
postdraws.cpool2_model1 <- inla.posterior.sample(100, completepool1_model1)
postdraws.cpool2_model2 <- inla.posterior.sample(100, completepool1_model2)

samples_1_1 <- data.frame()
for(j in 1:100){
  samples_1_1 <- rbind(samples_1_1, postdraws.cpool2_model1[[j]]$latent[104:110])
  colnames(samples_1) <- c("Intercept", "sat_2014")
}


beta0_sample_1 <- rnorm(100, mean=0, sd = 1)
beta1_sample_1 <- rnorm(100, mean=0, sd = 1)
tau1_sample_1 <- rgamma(100, shape = 0.01, scale = 0.01)

y_1_simulation <- beta0_sample_1 + samples_1$sat_2014*beta1_sample_1 + samples_1$Intercept + tau1_sample_1
# Create a grid of histograms and density plots

samples_2 <- data.frame()
for(j in 1:100){
  samples_2 <- rbind(samples_2, postdraws.cpool1_model2[[j]]$latent[104:110])
  colnames(samples_2) <- c("Intercept", "sat_2014")
}

beta0_sample_2 <- rnorm(100, mean=0, sd = 1)
beta1_sample_2 <- rnorm(100, mean=0, sd = 1)
tau1_sample_2 <- rgamma(100, shape = 0.01, scale = 0.01)

y_2_simulation <- beta0_sample_2 + samples_2$sat_2014*beta1_sample_2 + samples_2$Intercept + tau1_sample_2

plot1 <- ggplot(data.frame(y_1_simulation), aes(x=y_1_simulation)) +
  geom_histogram(aes(y=..density..)) +
  geom_density() +
  theme_bw()

plot2 <- ggplot(data.frame(y_2_simulation), aes(x=y_2_simulation)) +
  geom_histogram(aes(y=..density..)) +
  geom_density() +
  theme_bw()


plot_grid(plot1, plot2, ncol = 2)

```
The simulated data appears to to near about the same as the actual data shown in the Q2.2 histograms. Once we multiply the predictors with the hyperparameter estimates, we roughly get a similar histogram for both the models.  
# Question 4.1:

```{r}
library(sf)
library(tmap)
nc_sf <- st_read(system.file("shape/nc.shp", package="sf"),
quiet=TRUE)
st_crs(nc_sf)

global_rate <- sum(nc_sf$SID74)/sum(nc_sf$BIR74)
nc_sf$Expected <- global_rate * nc_sf$BIR74

tm_shape(nc_sf) + tm_fill("SID74")
tm_shape(nc_sf) + tm_fill("Expected")
```


```{r}
neighbor.nc <- poly2nb(nc_sf)
data(pennLC)
Wmat <- matrix(0, nrow=100, ncol=100)
for(m in 1:100){
  Wmat[m,neighbor.nc[[m]]] <- 1
  Wmat[neighbor.nc[[m]],m] <- 1
}  
Dmat <- diag(sapply(1:100, function(i) length(neighbor.nc[[i]])))

### Prior set 1
beta0.pr1 <- rnorm(n=100, mean=0, sd=1)
beta1.pr1 <- rnorm(n=100, mean=0, sd=1)
tauv.pr1 <- rgamma(n=100, shape=1, scale=1)
tauu.pr1 <- rgamma(n=100, shape=1, scale=1)


## Prior predictive 1
unstruc.re.priorpred1 <- matrix(NA, nrow=100, ncol=100)
spatial.re.priorpred1 <- matrix(NA, nrow=100, ncol=100)
lambda.priorpred1 <- matrix(NA, nrow=100, ncol=100)
ypriorpred1 <- matrix(NA, nrow=100, ncol=100)

linear.priorpred1 <- beta0.pr1 + beta1.pr1*matrix(nc_sf$Expected, nrow=100, ncol=100) 

for(j in 1:100){
  
  ## Unstructured Random Effect
  unstruc.re.priorpred1[,j] <- rnorm(n=100, mean=0, sd=1/sqrt(tauv.pr1[j])) 

  ## Structured Random Effect
  Qmat <- tauu.pr1[j]*(Dmat-Wmat) 
  Qmat.eigen <- eigen(Qmat)
  eigen.diag.comp <- 1/sqrt(Qmat.eigen$values)
  ### numerical instability
  eigen.diag.comp[is.na(eigen.diag.comp)] <- 0
  x <- Qmat.eigen$vectors%*%
    diag(eigen.diag.comp)%*%
    matrix(rnorm(n=100, mean=0, sd=1), nrow=100, ncol=1)
  spatial.re.priorpred1[,j] <- x  
 
  ## Lambda of Poisson
  lambda.priorpred1[,j] <- matrix(nc_sf$SID74, nrow=1, ncol=100)*
    exp(linear.priorpred1[,j] + 
          unstruc.re.priorpred1[,j] + 
          spatial.re.priorpred1[,j])

  ## Prior Predictive Counts
  ypriorpred1[,j] <- rpois(n = 100, lambda = lambda.priorpred1[,j])
}
```

```{r}

# Ashe, Alleghany, Surry, and Northampton

par(mfrow=c(2,2))

hist(ypriorpred1[1,], 
     main="Ashe: Prior predictive 1", 
     xlab = "counts",
     xlim=c(0, nc_sf$SID74[1] + 1500))
abline(v=nc_sf$SID74[1], lwd=3, col="red")
text(500, 20, paste("Number of NAs:", sum(is.na(ypriorpred1[1,]))))

hist(ypriorpred1[2,], 
     main="Alleghany: Prior predictive 1", 
     xlab = "counts",
     xlim=c(0, nc_sf$SID74[2] + 10))
abline(v=nc_sf$SID74[2], lwd=3, col="red")
text(5, 20, paste("Number of NAs:", sum(is.na(ypriorpred1[2,]))))

hist(ypriorpred1[3,], 
     main="Surry: Prior predictive 1", 
     xlab = "counts",
     xlim=c(0, nc_sf$SID74[3] + 20)
     )
abline(v=nc_sf$SID74[3], lwd=3, col="red")
text(15, 20, paste("Number of NAs:", sum(is.na(ypriorpred1[3,]))))

hist(ypriorpred1[5,], 
     main="Northampton: Prior predictive 1", 
     xlab = "counts", 
     xlim=c(0, nc_sf$SID74[5] + 100000))
abline(v=nc_sf$SID74[5], lwd=3, col="red")
text(50000, 20, paste("Number of NAs:", sum(is.na(ypriorpred1[5,]))))
```

# Question 4.2:

```{r}
## Values of E_i and neighborhood structure
counties <- nc_sf$CNTY_
neighbor.nc <- poly2nb(nc_sf)

prior.fixed2_model2 <- list(mean.intercept = 0, prec.intercept = 0.01,
                    mean = 0, prec = 0.01)
prior.prec2_model2 <- list(prec = list(prior = "loggamma", 
                               param = c(1, 1)))

nb2INLA("nc.adj", neighbor.nc)
g <- inla.read.graph(filename = "nc.adj")

nc_sf$re_u <- 1:nrow(nc_sf)
nc_sf$re_v <- 1:nrow(nc_sf)

formula1 <- deaths ~
  f(re_u, model = "besag", graph = g, 
    hyper = list(prec = list(prior = "loggamma",param = c(1, 1)))) + 
  f(re_v, model = "iid", 
    hyper = list(prec = list(prior = "loggamma",param = c(1, 1))))

#model1 <- inla(formula = formula1,
                            #data = data.frame(latcab),
                            #control.fixed = prior.fixed2_model2,
                            #control.family = list(hyper = list(prec = prior.prec2_model2)),
                            #control.compute=list(config = TRUE))
```


# Question 4.3:

```{r}
#nc_sf$popr1.1 <- ypostpred1[,1]

```