---
title: 'Lab Lecture #5'
output:
  html_document: default
  pdf_document: default
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, message = FALSE, warning = FALSE)
```


We will use the Pennsylvania data set we have already covered in class. More information on the data and an analysis can be found here: https://journal.r-project.org/archive/2018/RJ-2018-036/RJ-2018-036.pdf

```{r, message=FALSE, warning=FALSE}

library(SpatialEpi)
library(tidyverse)
library(sf)
library(spdep)

data(pennLC)
population <- pennLC$data$population
cases <- pennLC$data$cases
n.strata <- 16
E <- expected(population, cases, n.strata)
d <- aggregate(x = pennLC$data$cases, by = list(county = pennLC$data$county), FUN = sum)

# from spatial polygon to simple feature
pennLC.sf <- st_as_sf(pennLC$spatial.polygon)
pennLC.sf$county <- d$county
pennLC.sf$counts <- d$x
pennLC.sf$E <- E[match(pennLC.sf$county, unique(pennLC$data$county))]
pennLC.sf <- merge(pennLC.sf, pennLC$smoking, by.x = "county", by.y = "county")
pennLC.sf <- pennLC.sf%>%
  mutate(SIR = counts/E)

```

The BYM model for the data: 

$$
Y_i|\theta_i \sim Po(E_i  \theta_i) 
$$

$$
log(\theta_i) = \beta_0 + \beta_1 \times smoking_i + u_i + v_i
$$

$$
u_i|u_{j \sim i} \sim N\left( \frac{\sum_{j\sim i} u_j}{d_{i,i}}, \frac{1}{d_{i,i}\tau_u} \right)
$$
$$
v_i \sim N\left( 0, \frac{1}{\tau_v} \right)
$$

where $\beta_0$ is the intercept, $\beta_1$ relates to the effect of proportion of smokers on lung cancer, $\tau_u$ and $\tau_v$ are the precision of the spatial and unstructured random effect, respectively. The expected counts per county are denoted by $E_i$ and total number of neighbors for location $n_i$ is denoted as $d_{i,i}$. 

What is the geometry type and CRS of the Pennsylvania lip cancer data set (`pennLC.sf`)? Display counts, E, smoking, and SIR in a four panel plot. 

Geometry type: POLYGON

CRS: +proj=longlat +datum=WGS84 +no_defs 

```{r}

library(tmap)

tm_shape(pennLC.sf) + tm_polygons(c("counts", "E", "smoking", "SIR")) +
  tm_facets(ncol = 2)

```


We will fit a Besag-York-Mollié model to the data, but first do a few prior predictive checks. The ICAR structure is an improper and non-generative prior for the spatial random effect of the BYM model. However, we can impose a sum-to-zero constraint and use an eigenvalue decomposition to be able to generate data from the ICAR prior. 

Use the following two sets of priors to simulate 100 data sets from the prior predictive distribution of the BYM model: 

\begin{enumerate}
\item $\beta_0 \sim N(0, 1)$, $\beta_1 \sim N(0, 1)$, $\tau_v \sim Gamma(1, 1)$, $\tau_u \sim Gamma(1, 1)$ 
\item $\beta_0 \sim N(0, 10)$, $\beta_1 \sim N(0, 10)$, $\tau_v \sim Gamma(0.1, 0.1)$, $\tau_u \sim Gamma(0.1, 0.1)$ 
\end{enumerate}

For counties `centre`, `monroe`, `westmoreland`, `lebanon`, `beaver`, and `erie`, make a histogram of the prior predictive draws and add a vertical line at the observed value. 


One of the biggest issues with sampling from the ICAR is not only imposing the sum-to-zero constraint for identifiability, but also recognizing that eigen values _close_ to zero should also be set to zero. Depending on the approach taken, there may be a varied amount of numerical instability when sampling. 


```{r}

neighbor.penn <- poly2nb(pennLC.sf)
Wmat <- matrix(0, nrow=67, ncol=67)
for(m in 1:67){
  Wmat[m,neighbor.penn[[m]]] <- 1
  Wmat[neighbor.penn[[m]],m] <- 1
}  
Dmat <- diag(sapply(1:67, function(i) length(neighbor.penn[[i]])))

### Prior set 1
beta0.pr1 <- rnorm(n=100, mean=0, sd=1)
beta1.pr1 <- rnorm(n=100, mean=0, sd=1)
tauv.pr1 <- rgamma(n=100, shape=1, scale=1)
tauu.pr1 <- rgamma(n=100, shape=1, scale=1)

### Prior set 2
beta0.pr2 <- rnorm(n=100, mean=0, sd=10)
beta1.pr2 <- rnorm(n=100, mean=0, sd=10)
tauv.pr2 <- rgamma(n=100, shape=0.1, scale=0.1)
tauu.pr2 <- rgamma(n=100, shape=0.1, scale=0.1)

## Prior predictive 1
unstruc.re.priorpred1 <- matrix(NA, nrow=67, ncol=100)
spatial.re.priorpred1 <- matrix(NA, nrow=67, ncol=100)
lambda.priorpred1 <- matrix(NA, nrow=67, ncol=100)
ypriorpred1 <- matrix(NA, nrow=67, ncol=100)

linear.priorpred1 <- beta0.pr1 + 
  beta1.pr1*matrix(pennLC.sf$smoking, nrow=67, ncol=100) 

for(j in 1:100){
  
  ## Unstructured Random Effect
  unstruc.re.priorpred1[,j] <- rnorm(n=67, mean=0, sd=1/sqrt(tauv.pr1[j])) 

  ## Structured Random Effect
  Qmat <- tauu.pr1[j]*(Dmat-Wmat)
  print(Qmat)
  Qmat.eigen <- eigen(Qmat)
  eigen.diag.comp <- 1/sqrt(Qmat.eigen$values)
  ### numerical instability
  eigen.diag.comp[is.na(eigen.diag.comp)] <- 0
  x <- Qmat.eigen$vectors%*%
    diag(eigen.diag.comp)%*%
    matrix(rnorm(n=67, mean=0, sd=1), nrow=67, ncol=1)
  spatial.re.priorpred1[,j] <- x  
 
  ## Lambda of Poisson
  lambda.priorpred1[,j] <- matrix(pennLC.sf$E, nrow=1, ncol=67)*
    exp(linear.priorpred1[,j] + 
          unstruc.re.priorpred1[,j] + 
          spatial.re.priorpred1[,j])

  ## Prior Predictive Counts
  ypriorpred1[,j] <- rpois(n = 67, lambda = lambda.priorpred1[,j])
  
}

## Prior predictive 2  
unstruc.re.priorpred2 <- matrix(NA, nrow=67, ncol=100)
spatial.re.priorpred2 <- matrix(NA, nrow=67, ncol=100)
lambda.priorpred2 <- matrix(NA, nrow=67, ncol=100)
ypriorpred2 <- matrix(NA, nrow=67, ncol=100)

linear.priorpred2 <- beta0.pr2 + 
  beta1.pr2*matrix(pennLC.sf$smoking, nrow=67, ncol=100) 

for(j in 1:100){
  
  ## Unstructured Random Effect
  unstruc.re.priorpred2[,j] <- rnorm(n=67, mean=0, sd=1/sqrt(tauv.pr2[j])) 

  ## Structured Random Effect
  Qmat <- tauu.pr2[j]*(Dmat-Wmat) 
  Qmat.eigen <- eigen(Qmat)
  eigen.diag.comp <- 1/sqrt(Qmat.eigen$values)
  ### numerical instability
  eigen.diag.comp[is.na(eigen.diag.comp)] <- 0
  x <- Qmat.eigen$vectors%*%
    diag(eigen.diag.comp)%*%
    matrix(rnorm(n=67, mean=0, sd=1), nrow=67, ncol=1)
  spatial.re.priorpred2[,j] <- x  
 
  ## Lambda of Poisson
  lambda.priorpred2[,j] <- matrix(pennLC.sf$E, nrow=1, ncol=67)*
    exp(linear.priorpred2[,j] + 
          unstruc.re.priorpred2[,j] + 
          spatial.re.priorpred2[,j])

  ## Prior Predictive Counts
  ypriorpred2[,j] <- rpois(n = 67, lambda = lambda.priorpred2[,j])
  
}

```

Let's make the histograms for the prior predictive draws of `centre`, `monroe`, `westmoreland`, `lebanon`, `beaver`, and `erie`. 

```{r, fig.height=10}

counties <- c("centre", "monroe", "westmoreland", "lebanon", "beaver", "erie")

par(mfrow=c(6,1))

hist(ypriorpred1[4,], #xlim=c(0, pennLC.sf$counts[4] + 50),
     main="beaver: prior predictive 1", xlab = "counts")
abline(v=pennLC.sf$counts[4], lwd=3, col="red")
text(50, 20, paste("Number of NAs:", sum(is.na(ypriorpred1[4,]))))

hist(ypriorpred1[14,], #xlim=c(0, pennLC.sf$counts[14] + 50),
     main="centre: prior predictive 1" , xlab = "counts")
abline(v=pennLC.sf$counts[14], lwd=3, col="red")
text(50, 20, paste("Number of NAs:", sum(is.na(ypriorpred1[14,]))))

hist(ypriorpred1[25,], #xlim=c(0, pennLC.sf$counts[25] + 50),
     main="erie: prior predictive 1", xlab = "counts")
abline(v=pennLC.sf$counts[25], lwd=3, col="red")
text(50, 20, paste("Number of NAs:", sum(is.na(ypriorpred1[25,]))))

hist(ypriorpred1[38,], #xlim=c(0, pennLC.sf$counts[38] + 50),
     main="lebanon: prior predictive 1", xlab = "counts")
abline(v=pennLC.sf$counts[38], lwd=3, col="red")
text(50, 20, paste("Number of NAs:", sum(is.na(ypriorpred1[38,]))))

hist(ypriorpred1[45,], #xlim=c(0, pennLC.sf$counts[45] + 50),
     main="monroe: prior predictive 1", xlab = "counts")
abline(v=pennLC.sf$counts[45], lwd=3, col="red")
text(50, 20, paste("Number of NAs:", sum(is.na(ypriorpred1[45,]))))

hist(ypriorpred1[65,], #xlim=c(0, pennLC.sf$counts[65] + 50),
     main="westmoreland: prior predictive 1", xlab = "counts")
abline(v=pennLC.sf$counts[65], lwd=3, col="red")
text(50, 20, paste("Number of NAs:", sum(is.na(ypriorpred1[65,]))))

```


```{r, fig.height=10}

par(mfrow=c(6,1))

hist(ypriorpred2[4,], #xlim=c(0, pennLC.sf$counts[4] + 50),
     main="beaver: prior predictive 2", xlab = "counts")
abline(v=pennLC.sf$counts[4], lwd=3, col="red")
text(10, 20, paste("Number of NAs:", sum(is.na(ypriorpred2[4,]))))

hist(ypriorpred2[14,], #xlim=c(0, pennLC.sf$counts[14] + 50),
     main="centre: prior predictive 2" , xlab = "counts")
abline(v=pennLC.sf$counts[14], lwd=3, col="red")
text(10, 20, paste("Number of NAs:", sum(is.na(ypriorpred2[14,]))))

hist(ypriorpred2[25,], #xlim=c(0, pennLC.sf$counts[25] + 50),  
     main="erie: prior predictive 2", xlab = "counts")
abline(v=pennLC.sf$counts[25], lwd=3, col="red")
text(10, 20, paste("Number of NAs:", sum(is.na(ypriorpred2[25,]))))

hist(ypriorpred2[38,], #xlim=c(0, pennLC.sf$counts[38] + 50),
     main="lebanon: prior predictive 2", xlab = "counts")
abline(v=pennLC.sf$counts[38], lwd=3, col="red")
text(10, 20, paste("Number of NAs:", sum(is.na(ypriorpred2[38,]))))

hist(ypriorpred2[45,], #xlim=c(0, pennLC.sf$counts[45] + 50),
     main="monroe: prior predictive 2", xlab = "counts")
abline(v=pennLC.sf$counts[45], lwd=3, col="red")
text(10, 20, paste("Number of NAs:", sum(is.na(ypriorpred2[45,]))))

hist(ypriorpred2[65,], #xlim=c(0, pennLC.sf$counts[65] + 50), 
     main="westmoreland: prior predictive 2", xlab = "counts")
abline(v=pennLC.sf$counts[65], lwd=3, col="red")
text(10, 20, paste("Number of NAs:", sum(is.na(ypriorpred2[65,]))))

```


For prior sets 1 and 2 given in Question 1.2, create maps of 4 different prior predictive draws for lung cancer counts in Pennsylvania. You should have a total of 8 plots.  


```{r}

pennLC.sf$pp1.1 <- ypriorpred1[,1]
pennLC.sf$pp1.2 <- ypriorpred1[,2]
pennLC.sf$pp1.3 <- ypriorpred1[,3]
pennLC.sf$pp1.4 <- ypriorpred1[,4]

pennLC.sf$pp2.1 <- ypriorpred2[,1]
pennLC.sf$pp2.2 <- ypriorpred2[,2]
pennLC.sf$pp2.3 <- ypriorpred2[,3]
pennLC.sf$pp2.4 <- ypriorpred2[,4]

```

Maps -- no numbers in counties reflects NAs. 

```{r}

library(tmap)

## Maps of 0 and NAs
pp1.m1 <- tm_shape(pennLC.sf) + tm_polygons("pp1.1") + 
  tm_layout(main.title="Prior Predictive -- Prior 1") 
pp1.m2 <- tm_shape(pennLC.sf) + tm_polygons("pp1.2") +
  tm_layout(main.title="Prior Predictive -- Prior 1") 
pp1.m3 <- tm_shape(pennLC.sf) + tm_polygons("pp1.3") +
  tm_layout(main.title="Prior Predictive -- Prior 1") 
pp1.m4 <- tm_shape(pennLC.sf) + tm_polygons("pp1.4") + 
  tm_layout(main.title="Prior Predictive -- Prior 1") 

tmap_arrange(pp1.m1, pp1.m2, pp1.m3, pp1.m4)

## Maps of 0 and NAs
pp2.m1 <- tm_shape(pennLC.sf) + tm_polygons("pp2.1") +
  tm_layout(main.title="Prior Predictive -- Prior 2")
pp2.m2 <- tm_shape(pennLC.sf) + tm_polygons("pp2.2") +
  tm_layout(main.title="Prior Predictive -- Prior 2")
pp2.m3 <- tm_shape(pennLC.sf) + tm_polygons("pp2.3") +
  tm_layout(main.title="Prior Predictive -- Prior 2")
pp2.m4 <- tm_shape(pennLC.sf) + tm_polygons("pp2.4") +
  tm_layout(main.title="Prior Predictive -- Prior 2")

tmap_arrange(pp2.m1, pp2.m2, pp2.m3, pp2.m4)


```

Fit the model in INLA using the two different sets of priors and use the default priors specified in INLA (a total of 3 models). 

```{r, warning=FALSE, message=FALSE}

## Example code for the BYM model -- modify as needed
library(INLA)

## Values of E_i and neighborhood structure
E.penn <- pennLC.sf$E
neighbor.penn <- poly2nb(pennLC.sf)

nb2INLA("npenn.adj", neighbor.penn)
g <- inla.read.graph(filename = "npenn.adj")

pennLC.sf$re_u <- 1:nrow(pennLC.sf)
pennLC.sf$re_v <- 1:nrow(pennLC.sf)

```


```{r}
formula1 <- counts ~ smoking + 
  f(re_u, model = "besag", graph = g, 
    hyper = list(prec = list(prior = "loggamma",param = c(1, 1)))) + 
  f(re_v, model = "iid", 
    hyper = list(prec = list(prior = "loggamma",param = c(1, 1))))

res1 <- inla(formula1, family = "poisson", data = pennLC.sf, E = E.penn,
            control.predictor = list(compute = TRUE), 
            control.fixed = list(mean.intercept = 0, 
                                 prec.intercept = 1, 
                                 mean = 0, 
                                 prec = 1), 
            control.compute=list(config = TRUE))

formula2 <- counts ~ smoking + 
  f(re_u, model = "besag", graph = g, 
    hyper = list(prec = list(prior = "loggamma",param = c(0.1,  0.1)))) + 
  f(re_v, model = "iid", 
    hyper = list(prec = list(prior = "loggamma",param = c(0.1,  0.1))))

res2 <- inla(formula2, family = "poisson", data = pennLC.sf, E = E.penn,
            control.predictor = list(compute = TRUE), 
            control.fixed = list(mean.intercept = 0, 
                                 prec.intercept = 1/100, 
                                 mean = 0, 
                                 prec = 1/100), 
            control.compute=list(config = TRUE))

formula3 <- counts ~ smoking + 
  f(re_u, model = "besag", graph = g) + 
  f(re_v, model = "iid")

res3 <- inla(formula3, family = "poisson", data = pennLC.sf, E = E.penn,
            control.predictor = list(compute = TRUE),
            control.compute=list(config = TRUE))


summary(res1)
summary(res2)
summary(res3)

```

\begin{table}
\centering
\begin{tabular}{l c c c}
Model & Parameter & Mean & 95\% CI \\
\hline 
Model 1 & $\beta_0$ & `r res1$summary.fixed$mean[1]` & (`r res1$summary.fixed$"0.025quant"[1]`, `r res1$summary.fixed$"0.975quant"[1]`)\\
Model 1 & $\beta_1$ & `r res1$summary.fixed$mean[2]` & (`r res1$summary.fixed$"0.025quant"[2]`, `r res1$summary.fixed$"0.975quant"[2]`) \\
Model 1 & $\tau_u$ & `r res1$summary.hyperpar$mean[1]` &  (`r res1$summary.hyperpar$"0.025quant"[1]`, `r res1$summary.hyperpar$"0.975quant"[1]`)\\
Model 1 & $\tau_v$ & `r res1$summary.hyperpar$mean[2]` &  (`r res1$summary.hyperpar$"0.025quant"[2]`, `r res1$summary.hyperpar$"0.975quant"[2]`)\\
\hline
Model 2 & $\beta_0$ &  `r res2$summary.fixed$mean[1]` & (`r res2$summary.fixed$"0.025quant"[1]`, `r res2$summary.fixed$"0.975quant"[1]`)\\
Model 2 & $\beta_1$ &`r res2$summary.fixed$mean[2]`  &(`r res2$summary.fixed$"0.025quant"[2]`, `r res2$summary.fixed$"0.975quant"[2]`) \\
Model 2 & $\tau_u$ &`r res2$summary.hyperpar$mean[1]` & (`r res2$summary.hyperpar$"0.025quant"[1]`, `r res2$summary.hyperpar$"0.975quant"[1]`)\\
Model 2 & $\tau_v$ &`r res2$summary.hyperpar$mean[2]` & (`r res2$summary.hyperpar$"0.025quant"[2]`, `r res2$summary.hyperpar$"0.975quant"[2]`)\\
\hline
Model 3 & $\beta_0$ &`r res3$summary.fixed$mean[1]`  & (`r res3$summary.fixed$"0.025quant"[1]`, `r res3$summary.fixed$"0.975quant"[1]`)\\
Model 3 & $\beta_1$ & `r res3$summary.fixed$mean[2]` & (`r res3$summary.fixed$"0.025quant"[2]`, `r res3$summary.fixed$"0.975quant"[2]`)\\
Model 3 & $\tau_u$ &`r res3$summary.hyperpar$mean[1]` & (`r res3$summary.hyperpar$"0.025quant"[1]`, `r res3$summary.hyperpar$"0.975quant"[1]`)\\
Model 3 & $\tau_v$ &`r res3$summary.hyperpar$mean[2]` & (`r res3$summary.hyperpar$"0.025quant"[2]`, `r res3$summary.hyperpar$"0.975quant"[2]`)\\
\hline
\end{tabular}
\end{table}

Create two maps of the results for each fitted model. 

\begin{itemize}
\item Map 1: mean expected count of lung cancer cases across Pennsylvania. 
\item Map 2: standard deviation of expected counts of lung cancer cases across Pennsylvania
\end{itemize}


```{r, eval=FALSE}

## Example code for Map 1: 
res$summary.fitted.values[,"mean"]

## Example code for Map 2:
res$summary.fitted.values[,"sd"]

```

```{r}

pennLC.sf$model1.mean <- res1$summary.fitted.values[,"mean"]*E.penn
pennLC.sf$model1.sd <- res1$summary.fitted.values[,"sd"]*E.penn
pennLC.sf$model2.mean <- res2$summary.fitted.values[,"mean"]*E.penn
pennLC.sf$model2.sd <- res2$summary.fitted.values[,"sd"]*E.penn
pennLC.sf$model3.mean <- res3$summary.fitted.values[,"mean"]*E.penn
pennLC.sf$model3.sd <- res3$summary.fitted.values[,"sd"]*E.penn

```


Model 1

```{r}
res.m1.mean <- tm_shape(pennLC.sf) + 
  tm_polygons("model1.mean") + 
  tm_layout(main.title = "Predicted Counts")

res.m1.sd <- tm_shape(pennLC.sf) + 
  tm_polygons("model1.sd") + 
  tm_layout(main.title = "Standard Deviation of Predicted Counts")

tmap_arrange(res.m1.mean, res.m1.sd, ncol=1)
```

Model 2

```{r}
res.m2.mean <- tm_shape(pennLC.sf) + 
  tm_polygons("model2.mean") + 
  tm_layout(main.title = "Predicted Counts")

res.m2.sd <- tm_shape(pennLC.sf) + 
  tm_polygons("model2.sd") + 
  tm_layout(main.title = "Standard Deviation of Predicted Counts")

tmap_arrange(res.m2.mean, res.m2.sd, ncol=1)
```

Model 3

```{r}
res.m3.mean <- tm_shape(pennLC.sf) + 
  tm_polygons("model3.mean") + 
  tm_layout(main.title = "Predicted Counts")

res.m3.sd <- tm_shape(pennLC.sf) + 
  tm_polygons("model3.sd") + 
  tm_layout(main.title = "Standard Deviation of Predicted Counts")

tmap_arrange(res.m3.mean, res.m3.sd, ncol=1)

```

Simulate 100 data sets from the posterior predictive distribution per each of the three fitted models. For counties `centre`, `monroe`, `westmoreland`, `lebanon`, `beaver`, and `erie`, make a histogram of the posterior predictive draws and add a vertical line at the observed value. 

```{r}

postpred1 <- inla.posterior.sample(n=100, res1)
postpred2 <- inla.posterior.sample(n=100, res2)
postpred3 <- inla.posterior.sample(n=100, res3)

## Posterior predictive 1
unstruc.re.postpred1 <- matrix(NA, nrow=67, ncol=100)
spatial.re.postpred1 <- matrix(NA, nrow=67, ncol=100)
lambda.postpred1 <- matrix(NA, nrow=67, ncol=100)
ypostpred1 <- matrix(NA, nrow=67, ncol=100)
linear.postpred1 <- matrix(NA, nrow=67, ncol=100)

  unstruc.index <- which(rownames(postpred1[[1]]$latent) %in% paste("re_v:", 1:67, sep=""))
  struc.index <- which(rownames(postpred1[[1]]$latent) %in% paste("re_u:", 1:67, sep=""))
  intercept.index <- which(rownames(postpred1[[1]]$latent) %in% "(Intercept):1")
  slope.index <- which(rownames(postpred1[[1]]$latent) %in% "smoking:1")
    
for(j in 1:100){
  
  ## Unstructured Random Effect
  unstruc.re.postpred1[,j] <- postpred1[[j]]$latent[unstruc.index]

  ## Structured Random Effect
  spatial.re.postpred1[,j] <- postpred1[[j]]$latent[struc.index]  
 
  ## Linear Terms
  linear.postpred1[,j] <- postpred1[[j]]$latent[intercept.index] + 
  postpred1[[j]]$latent[slope.index]*pennLC.sf$smoking
  
  ## Lambda of Poisson
  lambda.postpred1[,j] <- pennLC.sf$E*
    exp(linear.postpred1[,j] + 
          unstruc.re.postpred1[,j] + 
          spatial.re.postpred1[,j])

  ## Prior Predictive Counts
  ypostpred1[,j] <- rpois(n = 67, lambda = lambda.postpred1[,j])
  
}


## Posterior predictive 2
unstruc.re.postpred2 <- matrix(NA, nrow=67, ncol=100)
spatial.re.postpred2 <- matrix(NA, nrow=67, ncol=100)
lambda.postpred2 <- matrix(NA, nrow=67, ncol=100)
ypostpred2 <- matrix(NA, nrow=67, ncol=100)
linear.postpred2 <- matrix(NA, nrow=67, ncol=100)
    
for(j in 1:100){
  
  ## Unstructured Random Effect
  unstruc.re.postpred2[,j] <- postpred2[[j]]$latent[unstruc.index]

  ## Structured Random Effect
  spatial.re.postpred2[,j] <- postpred2[[j]]$latent[struc.index]  
 
  ## Linear Terms
  linear.postpred2[,j] <- postpred2[[j]]$latent[intercept.index] + 
  postpred2[[j]]$latent[slope.index]*pennLC.sf$smoking
  
  ## Lambda of Poisson
  lambda.postpred2[,j] <- pennLC.sf$E*
    exp(linear.postpred2[,j] + 
          unstruc.re.postpred2[,j] + 
          spatial.re.postpred2[,j])

  ## Prior Predictive Counts
  ypostpred2[,j] <- rpois(n = 67, lambda = lambda.postpred2[,j])
  
}

## Posterior predictive 3
unstruc.re.postpred3 <- matrix(NA, nrow=67, ncol=100)
spatial.re.postpred3 <- matrix(NA, nrow=67, ncol=100)
lambda.postpred3 <- matrix(NA, nrow=67, ncol=100)
ypostpred3 <- matrix(NA, nrow=67, ncol=100)
linear.postpred3 <- matrix(NA, nrow=67, ncol=100)
    
for(j in 1:100){
  
  ## Unstructured Random Effect
  unstruc.re.postpred3[,j] <- postpred3[[j]]$latent[unstruc.index]

  ## Structured Random Effect
  spatial.re.postpred3[,j] <- postpred3[[j]]$latent[struc.index]  
 
  ## Linear Terms
  linear.postpred3[,j] <- postpred3[[j]]$latent[intercept.index] + 
  postpred3[[j]]$latent[slope.index]*pennLC.sf$smoking
  
  ## Lambda of Poisson
  lambda.postpred3[,j] <- pennLC.sf$E*
    exp(linear.postpred3[,j] + 
          unstruc.re.postpred3[,j] + 
          spatial.re.postpred3[,j])

  ## Prior Predictive Counts
  ypostpred3[,j] <- rpois(n = 67, lambda = lambda.postpred3[,j])
  
}
    
  
```

__Results for posterior predictive 1__

```{r, fig.height=10}

par(mfrow=c(6,1))

hist(ypostpred1[4,], xlim=c(0, pennLC.sf$counts[4]+50), 
     main="beaver: posterior predictive 1", xlab = "counts")
abline(v=pennLC.sf$counts[4], lwd=3, col="red")

hist(ypostpred1[14,], xlim=c(0, pennLC.sf$counts[14]+50), 
     main="centre: posterior predictive 1" , xlab = "counts")
abline(v=pennLC.sf$counts[14], lwd=3, col="red")

hist(ypostpred1[25,], xlim=c(0, pennLC.sf$counts[25]+50), 
     main="erie: posterior predictive 1", xlab = "counts")
abline(v=pennLC.sf$counts[25], lwd=3, col="red")

hist(ypostpred1[38,], xlim=c(0, pennLC.sf$counts[38]+50), 
     main="lebanon: posterior predictive 1", xlab = "counts")
abline(v=pennLC.sf$counts[38], lwd=3, col="red")

hist(ypostpred1[45,], xlim=c(0, pennLC.sf$counts[45]+50), 
     main="monroe: posterior predictive 1", xlab = "counts")
abline(v=pennLC.sf$counts[45], lwd=3, col="red")

hist(ypostpred1[65,], xlim=c(0, pennLC.sf$counts[65]+50), 
     main="westmoreland: posterior predictive 1", xlab = "counts")
abline(v=pennLC.sf$counts[65], lwd=3, col="red")
```


__Results for posterior predictive 2__

```{r, fig.height=10}

par(mfrow=c(6,1))

hist(ypostpred2[4,], xlim=c(0, pennLC.sf$counts[4]+50), 
     main="beaver: posterior predictive 2", xlab = "counts")
abline(v=pennLC.sf$counts[4], lwd=3, col="red")

hist(ypostpred2[14,], xlim=c(0, pennLC.sf$counts[14]+50), 
     main="centre: posterior predictive 2" , xlab = "counts")
abline(v=pennLC.sf$counts[14], lwd=3, col="red")

hist(ypostpred2[25,], xlim=c(0, pennLC.sf$counts[25]+50), 
     main="erie: posterior predictive 2", xlab = "counts")
abline(v=pennLC.sf$counts[25], lwd=3, col="red")

hist(ypostpred2[38,], xlim=c(0, pennLC.sf$counts[38]+50), 
     main="lebanon: posterior predictive 2", xlab = "counts")
abline(v=pennLC.sf$counts[38], lwd=3, col="red")

hist(ypostpred2[45,], xlim=c(0, pennLC.sf$counts[45]+50), 
     main="monroe: posterior predictive 2", xlab = "counts")
abline(v=pennLC.sf$counts[45], lwd=3, col="red")

hist(ypostpred2[65,], xlim=c(0, pennLC.sf$counts[65]+50), 
     main="westmoreland: posterior predictive 2", xlab = "counts")
abline(v=pennLC.sf$counts[65], lwd=3, col="red")
```

__Results for posterior predictive 3__

```{r, fig.height=10}

par(mfrow=c(6,1))

hist(ypostpred3[4,], xlim=c(0, pennLC.sf$counts[4]+50), 
     main="beaver: posterior predictive 3", xlab = "counts")
abline(v=pennLC.sf$counts[4], lwd=3, col="red")

hist(ypostpred3[14,], xlim=c(0, pennLC.sf$counts[14]+50), 
     main="centre: posterior predictive 3" , xlab = "counts")
abline(v=pennLC.sf$counts[14], lwd=3, col="red")

hist(ypostpred3[25,], xlim=c(0, pennLC.sf$counts[25]+50), 
     main="erie: posterior predictive 3", xlab = "counts")
abline(v=pennLC.sf$counts[25], lwd=3, col="red")

hist(ypostpred3[38,], xlim=c(0, pennLC.sf$counts[38]+50), 
     main="lebanon: posterior predictive 3", xlab = "counts")
abline(v=pennLC.sf$counts[38], lwd=3, col="red")

hist(ypostpred3[45,], xlim=c(0, pennLC.sf$counts[45]+50), 
     main="monroe: posterior predictive 3", xlab = "counts")
abline(v=pennLC.sf$counts[45], lwd=3, col="red")

hist(ypostpred3[65,], xlim=c(0, pennLC.sf$counts[65]+50), 
     main="westmoreland: posterior predictive 3", xlab = "counts")
abline(v=pennLC.sf$counts[65], lwd=3, col="red")
```

Per each of the three fitted models in Question 1.4, create maps of 4 different posterior predictive draws for lung cancer counts in Pennsylvania (a total of 12 maps). 

Adding columns of posterior predictive draws to `pennLC.sf`. 

```{r}

pennLC.sf$popr1.1 <- ypostpred1[,1]
pennLC.sf$popr1.2 <- ypostpred1[,2]
pennLC.sf$popr1.3 <- ypostpred1[,3]
pennLC.sf$popr1.4 <- ypostpred1[,4]

pennLC.sf$popr2.1 <- ypostpred2[,1]
pennLC.sf$popr2.2 <- ypostpred2[,2]
pennLC.sf$popr2.3 <- ypostpred2[,3]
pennLC.sf$popr2.4 <- ypostpred2[,4]

pennLC.sf$popr3.1 <- ypostpred3[,1]
pennLC.sf$popr3.2 <- ypostpred3[,2]
pennLC.sf$popr3.3 <- ypostpred3[,3]
pennLC.sf$popr3.4 <- ypostpred3[,4]

```

Making maps. 

```{r, fig.height=10}

library(tmap)

## Posterior predictive 1
popr1.m1 <- tm_shape(pennLC.sf) + tm_polygons(col="popr1.1") + 
  tm_layout(main.title="PosteriorPredictive -- Prior 1")
popr1.m2 <- tm_shape(pennLC.sf) + tm_polygons(col="popr1.2") +
  tm_layout(main.title="Posterior Predictive -- Prior 1")
popr1.m3 <- tm_shape(pennLC.sf) + tm_polygons(col="popr1.3") +
  tm_layout(main.title="Posterior Predictive -- Prior 1")
popr1.m4 <- tm_shape(pennLC.sf) + tm_polygons(col="popr1.4") + 
  tm_layout(main.title="Posterior Predictive -- Prior 1") 

tmap_arrange(popr1.m1, popr1.m2, popr1.m3, popr1.m4, ncol=1)

## Posterior predictive 2
popr2.m1 <- tm_shape(pennLC.sf) + tm_polygons(col="popr2.1") +
  tm_layout(main.title="Posterior Predictive -- Prior 2")
popr2.m2 <- tm_shape(pennLC.sf) + tm_polygons(col="popr2.2") +
  tm_layout(main.title="Posterior Predictive -- Prior 2") 
popr2.m3 <- tm_shape(pennLC.sf) + tm_polygons(col="popr2.3") +
  tm_layout(main.title="PosteriorPredictive -- Prior 2") 
popr2.m4 <- tm_shape(pennLC.sf) + tm_polygons(col="popr2.4") +
  tm_layout(main.title="Posterior Predictive -- Prior 2") 

tmap_arrange(popr2.m1, popr2.m2, popr2.m3, popr2.m4, ncol=1)

## Posterior predictive 3
popr3.m1 <- tm_shape(pennLC.sf) + tm_polygons(col="popr3.1") +
  tm_layout(main.title="Posterior Predictive -- Default Prior") 
popr3.m2 <- tm_shape(pennLC.sf) + tm_polygons(col="popr3.2") +
  tm_layout(main.title="Posterior Predictive -- Default Prior") 
popr3.m3 <- tm_shape(pennLC.sf) + tm_polygons(col="popr3.3") +
  tm_layout(main.title="Posterior Predictive -- Default Prior") 
popr3.m4 <- tm_shape(pennLC.sf) + tm_polygons(col="popr3.4") +
  tm_layout(main.title="Posterior Predictive -- Default Prior") 

tmap_arrange(popr3.m1, popr3.m2, popr3.m3, popr3.m4, ncol=1)

```


